---
title: "Econometrics II TA Session #8"
author: "Hiroki Kato"
output:
  pdf_document:
    latex_engine: xelatex
    md_extensions: +raw_attribute
    number_sections: true
    extra_dependencies: ["xcolor"]
    keep_tex: yes
fontsize: 12pt
header-includes:
  - \usepackage{zxjatype}
  - \setCJKmainfont[BoldFont = IPAゴシック]{IPA明朝}
  - \setCJKsansfont{IPAゴシック}
  - \setCJKmonofont{IPAゴシック}
  - \parindent = 1em
  - \newcommand{\argmax}{\mathop{\rm arg~max}\limits}
  - \newcommand{\argmin}{\mathop{\rm arg~min}\limits}
  - \DeclareMathOperator*{\plim}{plim}
---

```{r setup, include = FALSE, echo = FALSE, purl = FALSE}
# This is options to make pdf file. You should ignore
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = TRUE, 
  cache = FALSE,
  fig.pos = "h")
knitr::opts_knit$set(root.dir = "C:/Users/katoo/Desktop/2020EconometricsTA/R")
```


# Empirical Application of Panel Data Model: Earnings Equation

## Backgruond

A researcher wants to estimate the effect of full-time work experience on wages.
He uses a *balanced* panel of 595 individuals from 1976 to 1982, taken from the Panel Study of Income Dynamics (PSID).
The *balanced* panel data means that we can observe all individuals every year.

```{r data}
dt <- read.csv("./data/wages.csv")
head(dt, 14)
```

The variable `id` and `time` indicate individual and time indexs.
We use these two variables to apply panel data models.
Additionally, we use the following variables:

- `exp`: years of full-time work experience
- `sqexp`: squared value of `exp`
- `sex`: a dummy variable taking 1 if an individual is female
- `ed`: years of education
- `lwage`: logarithm of wage

```{r summary}
dt <- dt[,c("id", "time", "exp", "sex", "ed", "lwage")]
dt$sqexp <- dt$exp^2
dt$sex <- ifelse(as.character(dt$sex) == "female", 1, 0)
summary(dt)
```

## Pooled OLS

Using the OLS method, 
we want to estimate the following linear panel data model:

$$
  \text{lwage}_{it} = 
  \alpha + \beta_1 \cdot \text{exp}_{it} +
  \beta_2 \cdot \text{sqexp}_{it} +
  \beta_3 \cdot \text{sex}_{it} + 
  \beta_4 \cdot \text{ed}_{it} + u_{it}.
$$

We will discuss assumptions for applying the OLS method.
Let $\mathbf{X}_{it}$ be a $1 \times K$ (stochastic) explanatory vector.
This vector contains `exp`, `sqexp`, `sex` and `ed`.
Let $Y_{it}$ be a random variable of outcome, that is `lwage`.
The balanced panel data is given by

| | $i = 1$ | $i = 2$ | $\cdots$ | $i = n$ |
| :--: | :--: | :--: | :--: | :--: |
| $t = 1$ | $(Y_{11}, \mathbf{X}_{11})$ | $(Y_{21}, \mathbf{X}_{21})$ | $\cdots$ | $(Y_{n1}, \mathbf{X}_{n1})$ |
| $t = 2$ | $(Y_{12}, \mathbf{X}_{12})$ | $(Y_{22}, \mathbf{X}_{22})$ | $\cdots$ | $(Y_{n2}, \mathbf{X}_{n2})$ |
| $\vdots$ | $\vdots$ | $\vdots$ | $\cdots$ | $\vdots$ |
| $t = T$ | $(Y_{1T}, \mathbf{X}_{1T})$ | $(Y_{2T}, \mathbf{X}_{2T})$ | $\cdots$ | $(Y_{nT}, \mathbf{X}_{nT})$ |

Then, the linear panel data model can be rewritten as follows:

$$
  Y_{it} = \mathbf{X}_{it} \beta + u_{it}, \quad t = 1, \ldots, T, \quad i = 1, \ldots, n.
$$

Using notations $\underline{\mathbf{X}}_i = (\mathbf{X}'_{i1}, \ldots, \mathbf{X}'_{iT})'$
and $\underline{Y}_i = (Y_{i1}, \ldots, Y_{iT})'$, and $\underline{u}_i = (u_{i1}, \ldots, u_{iT})'$,
we can reformulate this model as follows:

$$
  \underline{Y}_i = \underline{\mathbf{X}}_i \beta + \underline{u}_i, \quad \forall i.
$$

Now, we assume 

1. $E[\mathbf{X}'_{it}u_{it}] = 0$, $\forall i, t$. This assumption, called *(contempraneous) exogneity assumption*, implies that $u_{it}$ and $\mathbf{X}_{it}$ are orthogonal in the conditional mean sence, $E[u_{it} | \mathbf{X}_{it}] = 0$. However, this assumption does not imply $u_{it}$ is uncorrelated with the explanatory variables in all time periods (strictly exogeneity), that is, $E[u_{it} | \mathbf{X}_{i1}, \ldots, \mathbf{X}_{iT}] = 0$. This assumption palces no restriction on the relationship between $\mathbf{X}_{is}$ and $u_{it}$ for $s\not=t$.
1. $E[\underline{\mathbf{X}}'_i\underline{\mathbf{X}}_i] \succ 0$.

Under these two assumptions,
the true parameter can be identified by 
$$
  \beta = E[\underline{\mathbf{X}}'_i\underline{\mathbf{X}}_i]^{-1} E[\underline{\mathbf{X}}'_i\underline{Y}_i].
$$
Hence, the OLSE (pooled OLSE) is given by 
$$
  \hat{\beta} 
  = \left( \frac{1}{n} \sum_{i=1}^n \underline{\mathbf{X}}'_i\underline{\mathbf{X}}_i \right)
  \left( \frac{1}{n} \sum_{i=1}^n \underline{\mathbf{X}}'_i\underline{Y}_i \right)
  = \left( \frac{1}{n} \sum_{i=1}^n \sum_{t=1}^T \mathbf{X}'_{it} \mathbf{X}_{it} \right)
  \left( \frac{1}{n} \sum_{i=1}^n \sum_{t=1}^T \mathbf{X}'_{it} Y_{it} \right).
$$

The pooled OLS estimator is consistent and asymptotically normally distributed.
$$
  \sqrt{n}(\hat{\beta} - \beta) \sim N(0, A^{-1} B A^{-1}),
$$
where $A = E[\underline{\mathbf{X}}'_i\underline{\mathbf{X}}_i]$ 
and $B = E[\underline{\mathbf{X}}'_i \underline{u}_i \underline{u}'_i \underline{\mathbf{X}}_i]$.
The consistent estimator of the asymptotic variance covariance matrix is given by 
$$
  \hat{A}^{-1} \hat{B} \hat{A}^{-1} = 
  \left( \frac{1}{n} \sum_{i=1}^n \underline{\mathbf{X}}'_i\underline{\mathbf{X}}_i \right)^{-1}
  \left( \frac{1}{n} \sum_{i=1}^n \underline{\mathbf{X}}'_i \underline{u}_i \underline{u}'_i \underline{\mathbf{X}}_i \right)
  \left( \frac{1}{n} \sum_{i=1}^n \underline{\mathbf{X}}'_i\underline{\mathbf{X}}_i \right)^{-1}
$$
The standard errors caclulated by this matrix is called *robust standard errors clustered by individuals*.

In `R`, the pooled OLSE can be obtained by `lm` function.
However, the `lm` function does not return the cluster-robust standard errors.
Thus, you need to caclulate them by yourself.
Here is a sample code.

```{r pooledOLS}
# OLSE
pool <- lm(lwage ~ exp + sqexp + sex + ed, data = dt)

# Clustered SE
X <- model.matrix(pool); uhat <- pool$residuals
uhatset <- matrix(0, nrow = nrow(X), ncol = nrow(X))

i_from <- 1; j_from <- 1
for (i in 1:max(dt$id)) {
  x <- as.numeric(rownames(dt))[dt$id == i]
  usq <- uhat[x] %*% t(uhat[x])
  i_to <- i_from + nrow(usq) - 1
  j_to <- j_from + ncol(usq) - 1
  uhatset[i_from:i_to, j_from:j_to] <- usq
  i_from <- i_to + 1; j_from <- j_to + 1
}

Ahat <- t(X) %*% X
Bhat <- t(X) %*% uhatset %*% X
clust_vcov <- solve(Ahat) %*% Bhat %*% solve(Ahat)
clust_se <- sqrt(diag(clust_vcov))

print("Pooled OLSE"); coef(pool)
print("SE of pooled OLSE"); clust_se
```


Alternatively, 
using the `plm` function (the package `plm`) and the `coeftest` function (the package `lmtest`),
you can obtain the asymptotic variance covariance matrix of pooled OLSE easily.
The `plm` function provides the panel data model.
When you want to estimate pooled OLS, you need to specify `model = "pooling"`.
Moreover, you should specify individual and time index using `index` augment.
This augment passes `index = c("individual index", "time index")`.
After estimating the pooled OLS by the `plm` function,
you must use the `coeftest` function to obtain the cluster-robust standard errors.
To caclulate the clustered standard errors,
you should use the `vcovHC` function in the `vcov` augment.

```{r pooledOLS_plm}
library(plm)
library(lmtest)
library(sandwich)
test <- plm(
  lwage ~ exp + sqexp + sex + ed, 
  data = dt, model = "pooling", index = c("id", "time"))
coeftest(test, vcov = vcovHC(test, type = "HC0", cluster = "group"))
```