---
title: "Econometrics II TA Session #8"
author: "Hiroki Kato"
output:
  pdf_document:
    latex_engine: xelatex
    md_extensions: +raw_attribute
    number_sections: true
    extra_dependencies: ["xcolor"]
    keep_tex: yes
fontsize: 12pt
header-includes:
  - \usepackage{zxjatype}
  - \setCJKmainfont[BoldFont = IPAゴシック]{IPA明朝}
  - \setCJKsansfont{IPAゴシック}
  - \setCJKmonofont{IPAゴシック}
  - \parindent = 1em
  - \newcommand{\argmax}{\mathop{\rm arg~max}\limits}
  - \newcommand{\argmin}{\mathop{\rm arg~min}\limits}
  - \DeclareMathOperator*{\plim}{plim}
---

```{r setup, include = FALSE, echo = FALSE, purl = FALSE}
# This is options to make pdf file. You should ignore
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = TRUE, 
  cache = FALSE,
  fig.pos = "h")
knitr::opts_knit$set(root.dir = "C:/Users/katoo/Desktop/2020EconometricsTA/R")
```


# Empirical Application of Panel Data Model: Earnings Equation

## Backgruond

A researcher wants to estimate the effect of full-time work experience on wages.
He uses a *balanced* panel of 595 individuals from 1976 to 1982, taken from the Panel Study of Income Dynamics (PSID).
The *balanced* panel data means that we can observe all individuals every year.

```{r data}
dt <- read.csv("./data/wages.csv")
head(dt, 14)
```

The variable `id` and `time` indicate individual and time indexs.
We use these two variables to apply panel data models.
Additionally, we use the following variables:

- `exp`: years of full-time work experience
- `sex`: an indicator of gender
- `ed`: years of education
- `lwage`: logarithm of wage

```{r summary}
dt <- dt[,c("id", "time", "exp", "sex", "ed", "lwage")]
summary(dt)
```