
# Empirical Application of Time Series Model: Nikkei 225 

## Background and Data

The "Nikkei225" is a stock price index published by Nihon Keizai Shimbun (hereafter, NIKKEI).
NIKKEI calculates this price index based on 225 high liquid brands listed with first section of the Tokyo Stock Exchange.
We use daily data of the Nikkei 225 index taken from the yahoo finance 
(<https://stocks.finance.yahoo.co.jp/stocks/detail/?code=998407.O>).
The time length is from January 4th 2019 to January 22 2021.
We have 498 observations.
We read a csv data which recodes the Nikkei 225 dairy index, using the `read.csv` function in `R`.
Since `R` recognize a time variable (e.g., `2021/01/22`) as a character string,
we need to define a time variable, using `as.Date()` function.
The data structure is as follows:

```{r nikkeidata}
library(tidyverse)
dt <- read.csv("data/nikkei225.csv", stringsAsFactor = FALSE)
dt$date <- as.Date(dt$date, format = "%Y/%m/%d")
head(dt)
```

There are five variables:

- `date`: date variable
- `open_price`: open price in day $t$
- `high_price`: high price in day $t$
- `low_price`: low price in day $t$
- `close_price`: closed price in day $t$

Mainly, we use the `date` and `close_price`.

```{r TimeSeriesPlot, echo = FALSE, fig.cap = "Time Series Data of Nikkei 225 (Closed Price)", out.width="90%", fig.align = 'center'}
ggplot(dt, aes(x = date, y = close_price)) + 
  geom_line(color = "blue") +
  geom_vline(aes(xintercept = as.Date("2019/12/31")), linetype = 2) +
  geom_vline(aes(xintercept = as.Date("2020/01/16")), linetype = 2) +
  geom_vline(aes(xintercept = as.Date("2020/11/07")), linetype = 2) +
  annotate(
    geom = "rect",
    xmin = as.Date("2020/04/07"), xmax = as.Date("2020/05/25"), 
    ymin = -Inf, ymax = Inf,
    fill = "grey80", alpha = 0.4
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2020/05/20"), y = 27500, 
    label = "First declaration of \n a state of emergency \n in Japan", 
    hjust = "center"
  ) +
  annotate(
    geom = "segment", 
    x = as.Date("2019/11/30"), y = 27500, 
    xend = as.Date("2019/12/31"), yend = 27500, 
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2019/11/29"), y = 27500, 
    label = "First case of Covid-19 in China", 
    hjust = "right"
  ) +
  annotate(
    geom = "segment", 
    x = as.Date("2019/11/30"), y = 26500, 
    xend = as.Date("2020/01/16"), yend = 26500, 
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2019/11/29"), y = 26500, 
    label = "First case of Covid-19 in Japan", 
    hjust = "right"
  ) +
  annotate(
    geom = "segment", 
    x = as.Date("2020/10/07"), y = 25500, 
    xend = as.Date("2020/11/07"), yend = 25500, 
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2020/10/06"), y = 25500, 
    label = "U.S. Presidential Election", 
    hjust = "right"
  ) +
  scale_x_date(date_breaks = "2 months", date_labels = "%y/%m") +
  labs(x = "Date (Year/Month)", y = "Closed price") +
  theme_minimal() +
  theme(
    # setting: background
    plot.background = element_rect(
      #fill="#87CEEB50", 
      color = "transparent"
    ),
    
    # setting: plot
    panel.border = element_rect(color = "white", fill = NA),
    panel.background = element_rect(fill = "white"),   
    panel.grid = element_line(color = "grey80", size = 0.1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    
    # setting: text
    plot.title = element_text(hjust=0.5,size=20),       
    plot.caption = element_text(size=8),       
    
    # setting: axis
    axis.text = element_text(color="black",size=8),    
    axis.title = element_text(size=10),
    axis.ticks.length = unit(0.25, "cm"),
    axis.ticks.x = element_line(size = 0.1),
    axis.ticks.y = element_line(size = 0.1),
    axis.line = element_line(size = 0.1),
    
    # setting: legend
    legend.title = element_text(size=8),               
    legend.text = element_text(size=8),                
    legend.key.size = unit(0.5,"cm"),
    legend.background = element_rect(color = "white"), 
    legend.position = "bottom"
  )
```

Figure \@ref(fig:TimeSeriesPlot) shows the time series of closed price of the Nikkei225.
We summarize some features as follows:

- After the COVID-19 occured in Japan and China, the Nikkei225 has drastically decreased.
- During the first declaration of a state of emergency in Japan, the Nikkei225's performance has been a V-shaped recovery.
- The Nikkei225 has sharply increased immediaterly before and after the U.S. presidential election.

Some may wonder if the negative shock of COVID-19 reflects the Nikkei225.
To discuss it, we need to consider following two potential concerns.

1. unlisted companies (such as small restaurant business) may suffer heavily from the negative shock of COVID-19.
2. the Nikkei225 does not represent a variation of price index of 225 brands. In principle, the Nikkei225 is a mathematical mean of stock price of 225 brands. In fact, the the stock prices of top five brands which contribute to the Nikkei225 have increased at 70\%. On the other hand, the stock price of other brands have decreased at 5\% from the begging of 2020 [^cite]. 

[^cite]: See <https://news.yahoo.co.jp/articles/f63a4627b298857a62ac329b1ed41a88c2721bd4>.

Anyway, we test the stationarity of this time series.

## Autoregressive of Order 1: AR(1) model

To check the stationarity of this time series, we consider the following AR(1) model:
$$ X_t = \beta X_{t-1} + \epsilon_t, $$
where $X_t$ is the closed price of Nikkei225 in day $t$.
We assume $(\epsilon_t)$ is a white noise process, $(\epsilon_t) \sim WN(0, \sigma^2)$. 
that is, $E(\epsilon_t) = 0$, $E(\epsilon_t^2) = \sigma^2$, and $Cov(\epsilon_t, \epsilon_{t+h}) = 0$ for $h \not= 0$.

To estimate the unkown parameters $\theta = (\beta, \sigma^2)$,
we use the maximum likelihood method.
We assume $\epsilon_t \sim N_{\mathbb{R}}N(0, \sigma^2)$ for inference purposes.
Then, the conditional log-likelihood is 
$$
  M_T(X_1, \ldots, X_T; \theta) 
  = \frac{1}{T} \sum_{t=2}^T \left\{ -\log(2\pi\sigma^2) - \frac{X_t - \beta X_{t-1}}{2\sigma^2} \right\}.
$$
The MLE $\hat{\theta}$ can be obtained by solving
$$ \hat{\theta} = \argmax_{\theta} M_T(X_1, \ldots, X_T; \theta). $$ 

In `R`, the library calles `astsa` provides a function to estimate AR(p) model, named `sarima()`.
Originally, this function estimates the (seasonal) ARIMA model, using the maximum likelihood method.
Note that the ARIMA$(p, q, d)$ model is 
$$ X_t - X_{t-d} = c + \sum_{i=1}^{p} \beta_i X_{t-i} + \sum_{i=1}^Q \theta_i \epsilon_{t-i}. $$
The second term corresponds to AR($p$), and 
the third term corresponds to MA($q$).
This function passes five augments,
a time-series data, autoregressive of order ($p$), moving average of order ($q$),
$d$-the difference series, and whether to include constatnt term $c$.
To estimate AR(1) model as described above,
we specify `p = 1`, `q = 0`, `d = 0`, and `no.constant = TRUE`.
The R snippet is as follows:

```{r AR1, eval = FALSE}
ar1 <- sarima(dt$close_price, p = 1, q = 0, d = 0, no.constant = TRUE)
```

```{r AR1 echo, include = FALSE, purl = FALSE}
ar1 <- sarima(dt$close_price, p = 1, q = 0, d = 0, no.constant = TRUE)
```

```{r result_AR1}
sprintf(
  "The estimated beta is %1.5f (s.e. = %1.5f)", 
  ar1$fit$coef, sqrt(ar1$fit$var.coef)
)
sprintf("The estimated squared sigma is %1.2f", ar1$fit$sigma2)
```


## Dickey-Fuller test

First, we derive the stationary condition in AR(1) model.
The $N$-time iterated substition of $X_{t-1}$ yields
$$ X_t = \beta^N X_{t-N} + \sum_{k = 0}^N \beta^k \epsilon_{t-k}. $$
If $|\beta| < 1$, then the first term of right hand side coverges to zero as $N \to \infty$.
Thus, under $|\beta| < 1$, the causal stationary solution is $X_t = \sum_{k=0}^{\infty} \beta^k \epsilon_{t-k}$.

Testing for stationarity is quivalent to test for $\beta = 1$ in AR(1).
The null hypothesis is that the process is **not** stationary.
The alternative hypothesis is that the process is stationary, that is, $\beta < 1$.
The Dickey-Fuller test provides this *one-sided* test.

To implement this test with `R`, 
we ues the package called `tseries`.
We use the function called `adf.test()` to carry out the Dickey-Fuller test.
We need to pass two augments in this function.
The first augment is time-series data.
The second one, named `k`, is the number of order ($p$).
In this example, we pass `k = 1`, that is, AR(1).

```{r DFtest}
library(tseries)
df1 <- adf.test(dt$close_price, k = 1)
sprintf("The DF stats is %1.4f (p-value = %1.4f)", df1$statistic, df1$p.value)
```

As a result, we cannot reject the null hypothesis.
Thus, the time series of closed price of Nikkei225 is not stationary
when we use data from January 4th 2020 to January 22 2021.

## Stationarity of Log Return of Nikkei225

We showed that the time series of closed price of Nikkei225 is not stationary.
Then, is the log return of Nikkei225 stationary?

To check it, we first construct the log return of Nikkei225, using the time series of closed price.
To construct log return variable, we use the `dplyr` library.
First, after reading dataset `dt`, we sort dataset by date using `arrange()` function.
Second, we make the lagged closed price, using `lag()` function.
This function picks up the variable contained in the previous `n`-the row.
Finally, we make the log return, using `mutate()` function which makes a new variable.
As a result, the data structure is as follows:

```{r ConstructLogReturn}
dt <- dt %>% 
  arrange(date) %>% 
  mutate(lag_close_price = dplyr::lag(close_price, n = 1)) %>% 
  mutate(log_return = log(close_price) - log(lag_close_price))
head(dt[,c("date", "close_price", "lag_close_price", "log_return")])
```

Figure \@ref(fig:TimeSeriesPlot2) plots the time series of the log return.
After the COVID-19 pandemic, the absolute value of log return is large.
However, the overall time series of this index seems to be stationary.

```{r TimeSeriesPlot2, echo = FALSE, fig.cap = "Time Series Data of Nikkei 225 (Log Return)", out.width="90%", fig.align = 'center'}
ggplot(dt, aes(x = date, y = log_return)) + 
  geom_line(color = "blue") +
  geom_vline(aes(xintercept = as.Date("2019/12/31")), linetype = 2) +
  geom_vline(aes(xintercept = as.Date("2020/01/16")), linetype = 2) +
  geom_vline(aes(xintercept = as.Date("2020/11/07")), linetype = 2) +
  annotate(
    geom = "rect",
    xmin = as.Date("2020/04/07"), xmax = as.Date("2020/05/25"), 
    ymin = -Inf, ymax = Inf,
    fill = "grey80", alpha = 0.4
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2020/05/20"), y = 0.08, 
    label = "First declaration of \n a state of emergency \n in Japan", 
    hjust = "center"
  ) +
  annotate(
    geom = "segment", 
    x = as.Date("2019/11/30"), y = 0.07, 
    xend = as.Date("2019/12/31"), yend = 0.07, 
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2019/11/29"), y = 0.07, 
    label = "First case of Covid-19 in China", 
    hjust = "right"
  ) +
  annotate(
    geom = "segment", 
    x = as.Date("2019/11/30"), y = 0.06, 
    xend = as.Date("2020/01/16"), yend = 0.06, 
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2019/11/29"), y = 0.06, 
    label = "First case of Covid-19 in Japan", 
    hjust = "right"
  ) +
  annotate(
    geom = "segment", 
    x = as.Date("2020/10/07"), y = 0.1, 
    xend = as.Date("2020/11/07"), yend = 0.1, 
    arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(
    geom = "text", 
    x = as.Date("2020/10/06"), y = 0.1, 
    label = "U.S. Presidential Election", 
    hjust = "right"
  ) +
  scale_x_date(date_breaks = "2 months", date_labels = "%y/%m") +
  labs(x = "Date (Year/Month)", y = "Log Return") +
  theme_minimal() +
  theme(
    # setting: background
    plot.background = element_rect(
      #fill="#87CEEB50", 
      color = "transparent"
    ),
    
    # setting: plot
    panel.border = element_rect(color = "white", fill = NA),
    panel.background = element_rect(fill = "white"),   
    panel.grid = element_line(color = "grey80", size = 0.1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    
    # setting: text
    plot.title = element_text(hjust=0.5,size=20),       
    plot.caption = element_text(size=8),       
    
    # setting: axis
    axis.text = element_text(color="black",size=8),    
    axis.title = element_text(size=10),
    axis.ticks.length = unit(0.25, "cm"),
    axis.ticks.x = element_line(size = 0.1),
    axis.ticks.y = element_line(size = 0.1),
    axis.line = element_line(size = 0.1),
    
    # setting: legend
    legend.title = element_text(size=8),               
    legend.text = element_text(size=8),                
    legend.key.size = unit(0.5,"cm"),
    legend.background = element_rect(color = "white"), 
    legend.position = "bottom"
  )
```

We estimate AR(1) model, using this time series data.

```{r AR2, eval = FALSE}
ar2 <- sarima(dt$log_return, p = 1, q = 0, d = 0, no.constant = TRUE)
```

```{r AR2 echo, include = FALSE, purl = FALSE}
ar2 <- sarima(dt$log_return, p = 1, q = 0, d = 0, no.constant = TRUE)
```

```{r result_AR2}
sprintf(
  "The estimated beta is %1.5f (s.e. = %1.5f)", 
  ar2$fit$coef, sqrt(ar2$fit$var.coef)
)
sprintf("The estimated squared sigma is %1.4f", ar2$fit$sigma2)
```

Using `adf.test()` function in the library `tseries`,
we statistically test the stationarity of time series of log return.

```{r DFtest2}
x <- dt[dt$date != as.Date("2019/01/04"), "log_return"]
df2 <- adf.test(x, k = 1)
sprintf("The DF stats is %1.4f (p-value = %1.4f)", df2$statistic, df2$p.value)
```

As a result, we can reject the null hypothesis.
This implies that the time series of the log return of Nikkei225 is stationary.

# Reference 
