---
title: "Econometrics II TA Session #5"
author: "Hiroki Kato"
output:
  pdf_document:
    latex_engine: xelatex
    md_extensions: +raw_attribute
    number_sections: true
    extra_dependencies: ["xcolor"]
    keep_tex: yes
fontsize: 12pt
header-includes:
  - \usepackage{zxjatype}
  - \setCJKmainfont[BoldFont = IPAゴシック]{IPA明朝}
  - \setCJKsansfont{IPAゴシック}
  - \setCJKmonofont{IPAゴシック}
  - \parindent = 1em
  - \newcommand{\argmax}{\mathop{\rm arg~max}\limits}
  - \newcommand{\argmin}{\mathop{\rm arg~min}\limits}
  - \DeclareMathOperator*{\plim}{plim}
---

```{r setup, include = FALSE, echo = FALSE, purl = FALSE}
# This is options to make pdf file. You should ignore
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  echo = TRUE, 
  cache = FALSE,
  fig.pos = "h")
knitr::opts_knit$set(root.dir = "C:/Users/katoo/Desktop/2020EconometricsTA/R")
```

# Empirical Application of Truncated Regression: Labor Participation of Married Women (1)

## Background and Data

To develop women's social advancement,
we should create environment to keep a good balance between work and childcare after marriage.
In this application, using the dataset of married women, 
we explore how much childcare prevents married women to participate in labor market.

Our dataset originally comes from Stata sample data. [^source3]
This dataset contains the following variables:

- `whrs`: Hours of work. This outcome variable is truncated from below at zero.
- `kl6`: the number of preschool children
- `k618`: The number of school‐aged children
- `wa`: age
- `we`: The number of years of education

[^source3]: <http://www.stata-press.com/data/r13/laborsub.dt>. Because this is dta file, we need to import it, using the `read.dta` function in the library `foreign`. I intentionally remove married women who could not participate in the labor market.

```{r laborData}
dt <- read.csv(file = "./data/labor.csv", header = TRUE,  sep = ",")
summary(dt)
```

## Model 

Since we cannot observe those who could not partiapte in the labor market (`whrs = 0`),
we use the truncated regression model.
Thus, the selection rule is as follows:

\begin{equation*}
  \begin{cases}
    y_i = x_i \beta + u_i &\text{if}\:\: s_i = 1  \\
    s_i = 1 &\text{if}\:\: a_1 < y_i < a_2
  \end{cases}.
\end{equation*}
where $u_i \sim N(0, \sigma^2)$.
In this case, $a_1 = 0$ and $a_2 = +\infty$.

Since we are interested in estimating $\beta$, we must condition on $s_i = 1$.
The probability density function of $y_i$ conditional on $(x_i, s_i = 1)$ is 

\begin{equation*}
  p_{\theta}(y_i | x_i, s_i = 1) = \frac{f(y_i | x_i)}{\int_0^{+\infty} f(y_i | x_i) dy_i}.
\end{equation*}
where $\theta = (\beta, \sigma^2)'$.
Because the distribution of $y_i$ depends on the distribution of $u_i$,
using $u_i = y_i - x_i \beta$, we obtain 

\begin{equation*}
  p_{\theta}(u_i | x_i, -x_i \beta < u_i) 
  = \frac{1}{\sigma} \frac{\phi(\frac{y_i - x_i \beta}{\sigma})}{1 - \Phi(\frac{- x_i \beta}{\sigma})}.
\end{equation*}
Thus, the log-likelihood function is 

\begin{equation*}
  M_n(\theta) 
  = \sum_{i=1}^n \log \left( \frac{1}{\sigma} \frac{\phi(\frac{y_i - x_i \beta}{\sigma})}{1 - \Phi(\frac{- x_i \beta}{\sigma})} \right).
\end{equation*}

We provide two ways to estimate truncated regression, using `R`.
First way is to define the log-likelihood function directly and minimize its function by `nlm` function.
Recall that `nlm` function provides the Newton method to minimize the function.
We need to give intial values in argument of this function.
Coefficients of explanatory variables, `b[3:6]`, are zero, 
and intercept, `b[2]`, and $\sigma$, `b[1]`, are given by mean and standard deviation of `whrs`, respectively.

```{r truncMle}
whrs <- dt$whrs
kl6 <- dt$kl6; k618 <- dt$k618
wa <- dt$wa; we <- dt$we

LnLik <- function(b) {
  sigma <- b[1]
  xb <- b[2] + b[3]*kl6 + b[4]*k618 + b[5]*wa + b[6]*we
  condp <- dnorm((whrs - xb)/sigma)/(1 - pnorm(-xb/sigma))
  LL_i <- log(condp/sigma)
  LL <- -sum(LL_i)
  return(LL)
}

init <- c(sd(whrs), mean(whrs), 0, 0, 0, 0)
est.LnLik <- nlm(LnLik, init, hessian = TRUE)
```

Second way is to use the function `truncreg` in the library `truncreg`.
This function must specify the trucated point in arguments `point` and `direction`.
If `direction = "left"`, the outcome variable is truncated from below at `point`, that is, `point < y`.
On the other hand, if `direction = "right"`, the outcome variable is truncated from above at `point`, that is, `y < point`.

```{r truncreg}
library(truncreg)
model <- whrs ~ kl6 + k618 + wa + we
est.trunc <- truncreg(model, data = dt, point = 0, direction = "left")
se.trunc <- sqrt(diag(vcov(est.trunc)))
```


## Interpretations

Table \ref{lfp} shows results of truncated regression estimated by two methods.
As a comparison, we also show the OLS result in column (3).
All specifications show that the number of preschool and school-aged children reduces the hours of work.
The size of coefficient of the number of preschool and school-aged children become stronger
when we apply the truncated regression.
Although the relationship between labor participation and women's characteristics is statistically insignificant,
size of coefficients largely differs among three specifications.

```{r tab_truncate, results = "asis"}
ols <- lm(model, data = dt)
coef.LnLik <- est.LnLik$estimate
se.LnLik <- sqrt(diag(solve(est.LnLik$hessian)))
names(coef.LnLik) <- c("sigma", names(coef(ols)))
names(se.LnLik) <- c("sigma", names(coef(ols)))

library(stargazer)
stargazer(
  ols, ols, ols,
  column.labels = c("Truncated (truncreg)", "Truncated (nlm)", "OLS"),
  coef = list(coef(est.trunc), coef.LnLik[2:6]),
  se = list(se.trunc, se.LnLik[2:6]),
  report = "vcs", keep.stat = c("n"),
  covariate.labels = c(
    "\\#.Preschool Children",
    "\\#.School-aged Children",
    "Age", "Education Years"
  ),
  add.lines = list(
    c("Estimated Sigma", 
      round(coef(est.trunc)[6], 3), round(coef.LnLik[1], 3)),
    c("Log-Likelihood", 
      round(est.trunc$logLik, 3), round(-est.LnLik$minimum, 3))
  ),
  omit.table.layout = "n", table.placement = "t",
  title = "Truncated Regression: Labor Market Participation of Married Women",
  label = "lfp",
  type = "latex", header = FALSE  
)
```

# Empirical Application of Tobit Regression: Labor Participation of Married Women (2)

## Background and Data

We continue to investigate the previous research question.
Our dataset comes from Stata sample data, which is same as the previous dataset.
Unlike the previous dataset, we now observe married woment who do not participate in the labor market.
Additionally, we introduce the new variable:

- `lfp`: a dummy variable taking 1 if observed unit works.

The previous dataset contains observations with `lfp = 1`.
In this application, we also use observations with `lfp = 0` to estimate tobit model.

```{r labor2Data}
dt <- read.csv(file = "./data/labor2.csv", header = TRUE,  sep = ",")
summary(dt)
```

## Model

Our dependent variable is censored from below at zero.
This is because married women chooses zero labor time 
if, without any constraint, the optimal labor time is negative (corner solution).
In this case, we should use the tobit model.
The tobit model is 

\begin{equation*}
  y_i = 
  \begin{cases}
    \mathbf{x}_i \beta + u_i &\mathrm{if}\:\: y_i > a  \\
    a                        &\mathrm{otherwise}
  \end{cases},
\end{equation*}
where $E(u_i) = 0$ and $\mathrm{Var}(u_i) = 0$.
In this application, we set $a = 0$.

Using this model, the probability of $y_i$ conditional on $x_i$ is defined by 

\begin{equation*}
  p_{\beta, \sigma^2}(y_i | x_i) = \mathbb{P}(y_i \le 0)^{1[y_i = 0]} f(y_i | \mathbf{x}_i)^{1 - 1[y_i = 0]}
\end{equation*}

where $f(y_i|x_i)$ is the probability density function conditional on $\mathbf{x}_i$,
$1[y_i = 0]$ is an indicator function returing 1 if $y_i = 0$.
Now, we assume the distribution $u_i | \mathbf{x}_i \sim N(0, \sigma^2)$.
Then, we can reformulate $\mathbb{P}(y_i \le 0)$ as follows:

\begin{equation*}
  \mathbb{P}(y_i \le 0) 
  = \mathbb{P}(-\mathbf{x}_i \beta \le u_i) 
  = \Phi \left( -\frac{\mathbf{x}_i \beta}{\sigma} \right)  
  = 1 - \Phi \left( \frac{\mathbf{x}_i \beta}{\sigma} \right),
\end{equation*}
where $\Phi(\cdot)$ is the cumulative distribution function of the stadnard normal distribution.
Note that the last equatility comes from symmetric property of the standard normal distribution.
Moreover, the density function $f$ is reformulated as follows:

\begin{equation*}
  f(y_i | \mathbf{x}_i) = \frac{1}{\sigma} \phi \left( \frac{y_i - \mathbf{x}_i \beta}{\sigma} \right).
\end{equation*}

Assuming iid sample, we obtain the join probability function as follows:

\begin{equation*}
  p_{\beta, \sigma^2}((y_i | x_i), i = 1, \ldots, n) 
  = \prod_{i=1}^n \left(1 - \Phi \left( \frac{\mathbf{x}_i \beta}{\sigma} \right) \right)^{1[y_i = 0]} 
  \left( \frac{1}{\sigma} \phi \left( \frac{y_i - \mathbf{x}_i \beta}{\sigma} \right) \right)^{1 - 1[y_i = 0]}.
\end{equation*}

We estimate $\log p_{\beta, \sigma^2}((y_i | x_i), i = 1, \ldots, n)$, using the maximum likelihood method.
In `R`, there are two ways to implement the tobit regression.
First way is to define the log-likelihood function directly and minimize its function by `nlm` function.
We need to give intial values in argument of this function.
As intial values, coefficients of explanatory variables, `b[3:6]`, are zero, 
and intercept, denoted by `b[2]`, and $\sigma$, denoted by `b[1]`, 
are given by mean and standard deviation of `whrs`, respectively.

```{r tobitMLE}
whrs <- dt$whrs
kl6 <- dt$kl6; k618 <- dt$k618
wa <- dt$wa; we <- dt$we

LnLik <- function(b) {
  sigma <- b[1]
  xb <- b[2] + b[3]*kl6 + b[4]*k618 + b[5]*wa + b[6]*we
  Ia <- ifelse(whrs == 0, 1, 0)
  F0 <- 1 - pnorm(xb/sigma)
  fa <- dnorm((whrs - xb)/sigma)/sigma
  LL_i <- Ia * log(F0) + (1 - Ia) * log(fa)
  LL <- -sum(LL_i)
  return(LL)
}

init <- c(sd(whrs), mean(whrs), 0, 0, 0, 0)
est.LnLik <- nlm(LnLik, init, hessian = TRUE)
coef.tobitNLM <- est.LnLik$estimate
se.tobitNLM <- sqrt(diag(solve(est.LnLik$hessian)))
```


Second way is to use the function `vglm` in the library `VGAM`.
First, we need to declare the tobit distribution (`tobit`), using the `family` augment.
The `tobit` function needs the censored point (the value of $a$) in arguments `Lower` and `Upper`.
When you specify `Lower`, the observed outcome is left-censored.
On the other hand, when you specify `Upper`, the observed outcome is right-censored.
In this application, we set `Lower = 0`.

```{r tobit}
library(VGAM)
tobitVGAM <- vglm(whrs ~ kl6 + k618 + wa + we, family = tobit(Lower = 0), data = dt)
coef.tobitVGAM <- coef(tobitVGAM)
coef.tobitVGAM[2] <- exp(coef.tobitVGAM[2])
se.tobitVGAM <- sqrt(diag(vcov(tobitVGAM)))[-2]
```


## Interpretations

Table \ref{lfp_tobit} shows results of tobit regression estimated by two methods.
As a comparison, we also show the OLS result in column (3).
Althourgh all specifications show the same sign of coefficients,
size of coefficients of censored regression becomes stronger than of OLS.
For example, the number of preschool and school-aged children reduces the hours of work.
The size of coefficient of the number of preschool and school-aged children become stronger
when we apply the censored regression.

```{r output_tobit, results = "asis"}
ols <- lm(whrs ~ kl6 + k618 + wa + we, data =dt)
names(coef.tobitNLM) <- c("sigma", names(coef(ols)))
names(se.tobitNLM) <- c("sigma", names(coef(ols)))
names(coef.tobitVGAM) <- c(names(coef(ols))[1], "sigma", names(coef(ols))[-1])
names(se.tobitVGAM) <- names(coef(ols))

stargazer(
  ols, ols, ols,
  column.labels = c("Tobit (vglm)", "Tobit (nlm)", "OLS"),
  coef = list(coef.tobitVGAM[-2], coef.tobitNLM[-1]),
  se = list(se.tobitVGAM, se.tobitNLM[-1]),
  report = "vcs", keep.stat = c("n"),
  covariate.labels = c(
    "\\#.Preschool Children",
    "\\#.School-aged Children",
    "Age", "Education Years"
  ),
  add.lines = list(
    c("Estimated Sigma", 
      round(coef.tobitVGAM[2], 3), round(coef.tobitNLM[1], 3)),
    c("Log-Likelihood", 
      round(logLik(tobitVGAM), 3), round(-est.LnLik$minimum, 3))
  ),
  omit.table.layout = "n", table.placement = "t",
  title = "Tobit Regression: Labor Market Participation of Married Women",
  label = "lfp_tobit",
  type = "latex", header = FALSE  
)
```